---
title: "New York City Squirrels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(tidyverse)

nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv") %>% 
    mutate(date = lubridate::mdy(date))
```


# Squirrels seen's activities by main color?


```{r}
nyc_squirrels %>% 
    
    # Wrangling
    group_by(primary_fur_color) %>% 
    summarise_at(vars(running:foraging), sum) %>% 
    ungroup() %>% 
    filter(!is.na(primary_fur_color)) %>% 
    pivot_longer(c(running:foraging), names_to = "activity") %>% 
    mutate(activity = str_to_title(activity)) %>% 
    mutate(activity = fct_reorder(activity, value)) %>% 
    
    # Visualization
    ggplot(aes(activity, value, fill = primary_fur_color)) +
    geom_col() +
    coord_flip() +
    labs(
        subtitle = "Squirrels seen's activities by main color",
        fill = "Main Color", 
        x = "", 
        y = "count"
    ) +
    scale_fill_manual(values = c("black", "#d2691e", "darkgray")) +
    theme(
        panel.background = element_rect(fill = "white"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank()
    )
```


# How is the distribution of squirrels throught the days of the week?


```{r}
nyc_squirrels %>% 
    
    # Wrangling
    count(date) %>% 
    mutate(weekday = lubridate::wday(date, label = TRUE, week_start=1)) %>% 
    group_by(weekday) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    
    # Visualization
    ggplot(aes(weekday, n)) +
    geom_col() +
    geom_label(aes(label = n)) +
    labs(
        title = "Seen's squirrels by day of week",
        subtitle = "Weekend rocks!",
        y = "Number of squirrels",
        x = ""
    )
```


# Does squirrels with same predominant color stay together?


```{r}
nyc_squirrels %>% 
    
    # Wrangling
    select(lon = long, lat, group = hectare, color = primary_fur_color) %>% 
    
    # Visualization
    ggplot(aes(lon, lat, fill = color)) +
    geom_polygon(aes(group = group)) +
    coord_quickmap() +
    scale_fill_manual(values = c("black", "#d2691e", "darkgray"))+
    theme(
        panel.background = element_rect(fill = "white"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank()
    )
```

# From David Robinson Screencast

```{r}
library(tidyverse)

nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")

nyc_squirrels %>% 
    count(zip_codes, sort = TRUE)
```

```{r}
nyc_squirrels %>% 
    ggplot(aes(long, lat)) +
    geom_point()
    
by_hectare <- nyc_squirrels %>% 
    group_by(hectare) %>% 
    summarize(long = mean(long),
              lat = mean(lat),
              pct_gray = mean(primary_fur_color == "Gray", na.rm = TRUE),
              n = n()) %>% 
    ungroup()

by_hectare %>% 
    filter(n >= 10) %>% 
    ggplot(aes(long, lat, size = n, color = pct_gray)) +
    geom_point() +
    theme_void()

by_hectare %>% 
    filter(n >= 10) %>% 
    ggplot(aes(lat, pct_gray)) +
    geom_point() +
    geom_smooth()

by_hectare %>% 
    mutate(n_gray = round(pct_gray * n)) %>% 
    glm(cbind(n_gray, n-n_gray) ~ lat, data = ., family = "binomial") %>% 
    summary()

```

Squirrels may be more likely to be gray the higher the north in the park you go.

```{r}
nyc_squirrels %>% 
    count(highlight_fur_color, sort = TRUE)
```


Does a squirrel run away?

```{r}
glm(runs_from ~ lat, data = nyc_squirrels, family = "binomial") %>% 
    summary()
```


Plotting with Central Park Shapefile - base to creata the shiny visualization

```{r}
library(sf)

central_park_sf <- read_sf("C:/Users/joaop/Documents/GitHub/tidytuesday/data_raw/central_park_shapefiles")

ggplot(central_park_sf) +
    geom_sf()

central_park_sf %>% 
    ggplot() +
    geom_sf(aes(color = footway))


by_hectare <- nyc_squirrels %>% 
    add_count(hectare) %>% 
    group_by(hectare, n) %>% 
    summarize_at(vars(long, lat, runs_from, indifferent, ends_with("ing")), mean) %>% 
    ungroup()


by_hectare %>% 
    filter(n >= 10) %>% 
    ggplot() +
    geom_sf(data = central_park_sf) +
    geom_point(aes(long, lat, size = n, color = runs_from)) +
    theme_void() +
    scale_color_gradient2(low = "blue", high = "red", mid = "pink", 
                          midpoint = .3, labels = scales::percent) +
    labs(color = "% of squirrels run", 
         size = "# of squirrels", 
         title = "Squirrels in the northwest corner of Central Park are more likely to run away") +
    coord_sf(datum = NA)
```

Now go to app.R in the nyc-squirrels-app folder





